Tracking the hazy boundary between resonant and normal Auger decay near the K-edges of aqueous \ki~and \cli

### 2018-02-21 master.tex
feb 19 (gunnar, andreas, christophe)
Gunnar's, Andreas' and Christophe's corrections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}

@@

We thank Prof.\ Nobuhiro Kosugi and Dr.\ Matja\v{z} \v{Z}itnik for the fruitful discussions. Experiments were performed at the GALAXIES beamline, SOLEIL Synchrotron, France (Proposal No. 20140160). The authors are grateful to the SOLEIL staff for assistance during the beamtime. This project has received funding from the Research Executive Agency (REA) under the European Union$'$s Horizon 2020 research and innovation programme Grant agreement No 705515. Campus France and the PHC SIAM exchange program are acknowledged for financial support. N. V. Kryzhevoi and L. S. Cederbaum acknowledge the financial support of the European Research Council (ERC) (Advanced Investigator Grant No. 692657) and the Deutsche Forschungsgemeinschaft (DFG research unit 1789).

@@

\begin{itemize}
  \item suppinfo.pdf: contains the radial density distributions of the core excited states of the bare ions.
\end{itemize}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title{The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}
\title{Tracking the hazy boundary between resonant and normal Auger decay near the K-edges of aqueous \ki~and \cli}

@@

We thank Prof.\ Nobuhiro Kosugi and Dr.\ Matja\v{z} \v{Z}itnik for the fruitful discussions. Experiments were performed at the GALAXIES beamline, SOLEIL Synchrotron, France (Proposal No. 20140160). The authors are grateful to the SOLEIL staff for assistance during the beamtime. This project has received funding from the Research Executive Agency (REA) under the European Union$'$s Horizon 2020 research and innovation programme Grant agreement No 705515. Campus France and the PHC SIAM exchange program are acknowledged for financial support (project No. 38282QB). N. V. Kryzhevoi and L. S. Cederbaum acknowledge the financial support of the European Research Council (ERC) (Advanced Investigator Grant No. 692657) and the Deutsche Forschungsgemeinschaft (DFG research unit 1789).

@@

\begin{itemize}
  \item suppinfo.pdf: contains 1) the radial density distributions of the core excited states of the bare ions; 2) partial cross sections  and charge transfer time extracted from the experimental 2D map near the Cl 1s edge.
\end{itemize}

### 2018-02-21 methods.tex
feb 19 (gunnar, andreas, christophe)
Gunnar's, Andreas' and Christophe's corrections

For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline \citep{ceolin13:188,rueff15:175}. A differentially-pumped tube in which the microjet head is inserted, is mounted on a 3-axes motorized manipulator in front of the spectrometer lens. Two holes of 2\,mm diameter allow the photons to go in and out. At the end of the tube and in front of the lens, a 500\,$\mu$m diameter hole skimmer allows the electrons created at the interaction point to go in the direction of the spectrometer. The microjet head is mostly composed of a 30\,$\mu$m diameter vertical glass capillary facing a temperature-controlled catcher in CuBe having a 300\,$\mu$m hole, and a camera. Piezo motors allow their precise alignment relative to each other and to the photon beam. The catcher is placed at a distance of about 5\,mm from the capillary and is permanently pumped in order to extract the liquid. For the present experiment, a 0.5M KCl aqueous solution is injected in the capillary by a high performance liquid chromatography (HPLC) pump with a constant flux of 1.6\,ml/min. The alignment of the setup is performed on the KCl aqueous solution by measuring the water O1s XPS peak intensity and by optimizing the liquid phase vs gas phase ratio. The pressure in the main chamber is kept below the 10$^{-5}$\,mbar range whereas it is kept at about 10$^{-4}$\,mbar in the differentially-pumped tube when the HPLC pump is ON. Our equipment is an updated version of the equipment used in the reference \cite{faubel88:269}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution. The spectrometer resolution of about 0.6\,eV was achieved with the 500\,eV pass energy and 0.5\,mm slits. The photon energy resolution achieved at 2.8\,keV and 3.6\,keV was about 0.3\,eV and 0.4\,eV, respectively. The experimental 2D maps representing the evolution of the KLL Auger spectra in the vicinity of the \cli and \ki~K-edges, as a function of the photon energy, are shown in Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}, respectively. The aqueous \ki~and \cli~1s ionization potentials were measured at h$\nu$=5keV and calibrated on the liquid contribution of the O1s XPS spectrum \cite{winter06:1176}. The maps were also calibrated using the O1s photoelectron line of liquid water but at photon energies close to the potassium and chloride 1s ionization thresholds.

>>>>

For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline \citep{ceolin13:188,rueff15:175}. A differentially-pumped tube in which the microjet head is inserted, is mounted on a 3-axes motorized manipulator in front of the spectrometer lens. Two holes of 2\,mm diameter allow the photons to go in and out. At the end of the tube and in front of the lens, a 500\,$\mu$m diameter hole skimmer allows the electrons created at the interaction point to go in the direction of the spectrometer. The microjet head is mostly composed of a 30\,$\mu$m diameter vertical glass capillary facing a temperature-controlled catcher in CuBe having a 300\,$\mu$m hole, and a camera. Piezo motors allow their precise alignment relative to each other and to the photon beam. The catcher is placed at a distance of about 5\,mm from the capillary and is permanently pumped in order to extract the liquid. For the present experiment, a 0.5M KCl aqueous solution is injected in the capillary by a high performance liquid chromatography (HPLC) pump with a constant flux of 1.6\,ml/min. The alignment of the setup is performed on the KCl aqueous solution by measuring the water O1s XPS peak intensity and by optimizing the liquid vs gas phase ratio. The pressure in the main chamber is kept below the 10$^{-5}$\,mbar range whereas it is kept at about 10$^{-4}$\,mbar in the differentially-pumped tube when the HPLC pump is ON. Our equipment is an updated version of the equipment used in Ref.\ \cite{faubel88:269}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution. The spectrometer resolution of about 0.6\,eV was achieved with the 500\,eV pass energy and 0.5\,mm slits. The photon energy resolution achieved at 2.8\,keV and 3.6\,keV was about 0.3\,eV and 0.4\,eV, respectively. The experimental 2D maps representing the evolution of the KLL Auger spectra in the vicinity of the \cli and \ki~K-edges, as a function of the photon energy, are shown in Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}, respectively. The aqueous \ki~and \cli~1s ionization potentials were measured at h$\nu$=5keV and calibrated on the liquid contribution of the O1s XPS spectrum \cite{winter06:1176}. The maps were also calibrated using the O1s photoelectron line of liquid water but at photon energies close to the potassium and chloride 1s ionization thresholds.

### 2018-02-21 results.tex
feb 19 (gunnar, andreas, christophe)
Gunnar's, Andreas' and Christophe's corrections

@@

Finally, the normal Auger $^1$D main line of K$^{+}$ differs from that of Cl$^{-}$ by the presence of a large shoulder (feature A on Fig.\ \ref{fg:2dmap_k}) on the low kinetic energy side at about 2965\,eV kinetic energy. This shoulder is attributed to electron transfer from the solvent water molecules to the 3d orbital of K$^{+}$ \citep{ceolin17:263003}. In the case of \cli, there is no experimental evidence of such intense electron transfer processes.

@@

Further we assume that only the lowest peak in the theoretical XAS spectra is populated in the experiment for two reasons. First, due to the lifetime broadening, it spreads over approximately 2\,eV which coincides with the width of the pre-edge structure in the experimental XAS spectra. Second, the splitting between the first core excited state and the ionization threshold in the experiment is 1.2\,eV for \ki~and 0.2\,eV for \cli, and thus it is smaller than the splitting between the first and second peak in the theoretical spectra (1.5\,eV for \ki~and $\sim$3\,eV for \cli, Fig.\ \ref{fg:xas_kcl}). In the 6-coordinated cluster (Fig.\ \ref{fg:xas_kcl} upper left panel), which represents the complete first solvation shell around \ki, the lowest peak in the spectrum contains three states. The lowest and highest lying states are split by approximately 0.5\,eV and they have mixed 4p and 5p character. The low intensity state in between these two states has a predominantly 1s$^{-1}$3d character. Since the dispersive feature B appears at lower excitation energies compared to the feature C, we assume that it is produced in the resonant Auger decay of the lowest core excited states of \ki, which are predominantly of 1s$^{-1}$4p character. Moreover, we can attribute the feature C to the resonant Auger decay of the low intensity dipole forbidden 1s$^{-1}$3d state. Thus, we explain both the energy splitting of $\sim$400\,meV photon energy of the two features, and the fact that island C has lower intensity than feature B.

@@

As mentioned above, the delocalization of core excited electrons in aqueous solutions is ultrafast and as such it competes with the resonant Auger decay. In order to estimate the delocalization rate of the core excited electron at the pre-edges of \ki~and \cli, we used the core-hole clock method as in references \cite{bjorneholm92:1892,karis96:1380}.

In the case of \cli, it was possible to perform the same data treatment as in Ref.\ \cite{ceolin15:022502}, i.e., for each photon energy step, all components of the 2D map shown in Fig.\ \ref{fg:2dmap_cl} were isolated by fitting procedures and their intensity integrated to get a partial electron yield as a function of the photon energy. The result is shown on Fig.\ \ref{fg:si_ct_time} in the SI. The figure shows that there is a large overlap between the resonant and normal Auger contributions, due to the proximity of the resonance to the ionization potential and due to the very short lifetime of the corresponding states. At the specific photon energy corresponding to the lowest core excitation, $h\nu = 2825.2$\,eV (Fig.\ \ref{fg:2dmap_cl}, upper panel) a double-peak structure is observed in the interval of kinetic energies 2380 -- 2385\,eV . The position of the first peak coincides with the $^1$D main line resulting from normal Auger decay, whereas the second peak at 2383.5\,eV corresponds to the resonant Auger decay to the 2p$^{-2}$($^1$D)4p states. By fitting this double-peak structure with two Voigt functions, we determine the ratio of the intensities of these peaks to be $l/d \approx 1$. From the ratio $l/d$ and the Auger lifetime $\tau_{\text{c}}$, one can determine the delocalization time $\tau_{\text{CT}}$ according to the relation $\tau_{\text{CT}} = \tau_{\text{c}}l/d$ \citep{foehlisch05:373}{\color{red}MORE REFS}. Consequently, the delocalization time $\tau_{\text{CT}}$ is of the same order as the Auger lifetime, i.e.\ $\sim$1\,fs. The fast delocalization in this case is a result of the fact that the energy splitting between the \cli~(1s$^{-1}$4p) resonance and the ionization threshold is 0.2\,eV, and thus, smaller than the lifetime broadening of 0.62\,eV \citep{ceolin17:263003}.

>>>>

{\color{red}Gunnar: Including this discussion will invite questions about the energy calibration, and I’m not sure how this was done in these cases. Photoinduced charging may introduce shifts, that can vary with cross section}

@@

Finally, the normal Auger $^1$D main line of K$^{+}$ differs from that of Cl$^{-}$ by the presence of a large shoulder (feature A on Fig.\ \ref{fg:2dmap_k}) on the low kinetic energy side at about 2965\,eV kinetic energy. This shoulder is attributed to electron transfer from the solvent water molecules to the unoccupied 3d orbitals of K$^{+}$ \citep{ceolin17:263003}. In the case of \cli, there is no experimental evidence of such intense electron transfer processes.

@@

Further we assume that only the lowest peak in the theoretical XAS spectra is populated in the experiment for two reasons. First, due to the lifetime broadening, it spreads over approximately 2\,eV which coincides with the width of the pre-edge structure in the experimental XAS spectra. Second, the splitting between the first core excited state and the ionization threshold in the experiment is 1.2\,eV for \ki~and 0.2\,eV for \cli, and thus it is smaller than the splitting between the first and second peak in the theoretical spectra (1.5\,eV for \ki~and $\sim$3\,eV for \cli, Fig.\ \ref{fg:xas_kcl}). In the 6-coordinated cluster (Fig.\ \ref{fg:xas_kcl} upper left panel), which represents the complete first solvation shell around \ki, the lowest peak in the spectrum contains three states. The lowest and highest lying states are split by approximately 0.5\,eV and they have mixed 4p and 5p character. The low intensity state in between these two states has a predominantly 1s$^{-1}$3d character. Since the dispersive feature B appears at lower excitation energies compared to the feature C, we assume that it is produced in the resonant Auger decay of the lowest core excited states of \ki, which are predominantly of 1s$^{-1}$4p character. Moreover, we can attribute the feature C to the resonant Auger decay of the low intensity dipole forbidden 1s$^{-1}$3d state. Thus, we explain both the energy splitting of $\sim$400\,meV photon energy of the two features, and the fact that island C has lower intensity than feature B (Fig.\ \ref{fg:2dmap_k}).

@@

As mentioned above, the delocalization of core excited electrons in aqueous solutions is ultrafast and as such it competes with the resonant Auger decay. In order to estimate the delocalization rate of the core excited electron at the pre-edges of \ki~and \cli, we used the core-hole clock method \cite{bjorneholm92:1892,karis96:1380}.

In the case of \cli, it was possible to perform the same data treatment as in Ref.\ \cite{ceolin15:022502} i.e., for each photon energy step, all components of the 2D map shown in Fig.\ \ref{fg:2dmap_cl} were isolated by fitting procedures and their intensity integrated to get a partial electron yield as a function of the photon energy. The result is shown on Fig.\ \ref{fg:si_ct_time} in the SI. The figure shows that there is a large overlap between the resonant and normal Auger contributions, due to the proximity of the resonance to the ionization potential and due to the very short lifetime of the corresponding states. At the specific photon energy corresponding to the lowest core excitation, $h\nu = 2825.2$\,eV (Fig.\ \ref{fg:2dmap_cl}, upper panel) a double-peak structure is observed in the interval of kinetic energies 2380 -- 2385\,eV . The position of the first peak coincides with the $^1$D main line resulting from normal Auger decay, whereas the second peak at 2383.5\,eV corresponds to the resonant Auger decay to the 2p$^{-2}$($^1$D)4p states. By fitting this double-peak structure with two Voigt functions, we determine the ratio of the intensities of these peaks to be $l/d \approx 1$. From the ratio $l/d$ and the Auger lifetime $\tau_{\text{c}}$, one can determine the delocalization time $\tau_{\text{CT}}$ according to the relation $\tau_{\text{CT}} = \tau_{\text{c}}l/d$ \citep{foehlisch05:373}{\color{red}MORE REFS}. Consequently, the delocalization time $\tau_{\text{CT}}$ is of the same order as the Auger lifetime, i.e.\ $\sim$1\,fs. The fast delocalization in this case is a result of the fact that the energy splitting between the \cli~(1s$^{-1}$4p) resonance and the ionization threshold is 0.2\,eV, and thus, smaller than the lifetime broadening of 0.62\,eV \citep{ceolin17:263003}.
